Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$: limitations imposed by cosmic variance

### 2013-07-30 paper/ms.tex
new title, different abstract and improved intro

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Constraints on the mass and occupation
  fraction of dark matter halos hosting Lyman-alpha emitters
  at $z\sim 3$}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo with mass in the range $M_{\rm
  min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We define three different families of
parameters consistent with observations based on their mass range
$\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ dex and occupation
fraction $f_{\rm occ}$.  From this we conclude that the number
density and spatial clustering information available from observations
is not sufficient to derive a unique and narrow set of values for the
parameters in the model. In particular, it is always possible to find a 
a model with any desired occupation fraction in the range $0.1\leq
f_{\rm occ}\leq 1.0$.  However, there is a dominant family composed by
models with a narrow mass range $\Delta M<1.0$ dex, a low occupation fraction
$f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm   max}<
10^{12}$\hMsun.  The existence of this dominant family gives support
to the idea that the most massive dark matter halos at that epoch do
not host the brightest LAEs and that only a small fraction of star
forming galaxies can be actually detected as LAEs.

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations, markers of the galaxy formation process through
cosmic history \citep{ForeroRomero2012}. and tracers of active star formation 
\citep{Guaita2013}

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter, whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980}

Most models of galaxy formation find that the mass of the halo can be
used to predict properties of the galaxy such as its stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
mass. Furthermore, the spatial clustering of galaxies on large scales
is entirely dictated by the halo distribution.  For the reasons
mentioned above, finding the typical dark matter halo mass hosting
LAEs represents a significant step forward to understand the nature of
this population in the context of Lambda Cold Dark Matter
($\Lambda$CDM) paradigm.

@@

Added to the uncertainties in the astrophysical processes describing
star formation in galactic populations, a highly debated step in this
approach is the calculation of the fraction of \ly photons
that escape the galaxy to the observer. Given the resonant nature of
the \lya line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM) \citep{Neufeld1991, ForeroRomero2011, Dijkstra2012,
Laursen2013,Orsi2012}.

@@

Our objective in the present \documentname is to use a method to
bypass all the complexity involved in the estimation of the \ly escape
fraction to infer information relevant to the understanding of LAEs in
a cosmological context. The analysis we present is thus based on one
of the most robust prediction of $\Lambda$CDM, namely the clustering
and abudance of dark matter halos. We construct a simple model that
allow us to construct mock catalogs and explore a wide range of
parameters to narrow the mass interval of halos hosting LAEs together
with its occupation fraction.

Another central point in our study is the explicit inclusion of cosmic
variance effects by using large volume cosmological simulations. This
allows us to go closer to the space of observables and directly
explore the compatibility of a model with observations.

This paper is structured as follows. In the next section we present
the simulation and model used to produce the mock catalogs and the criteria
used to compare them against observations. In \SS \ref{sec:results} we
present the main results in terms of the parameters of our model. We
continue with a discussion of these results under the light of other
observational results for LAEs and abundance matching models for
galaxy formation before presenting our conclusions in \SS
\ref{sec:conclusions}.

>>>>

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$: limitations imposed by cosmic variance}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
Jaime E. Forero-Romero$^{1}$ and
Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching observational geometries. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host one LAE with a
probability $f_{\rm occ}$. We find that the large scatter in the mocks, induced by
cosmic variance, makes the number density and spatial clustering
information insufficient to derive a
unique and narrow set of values for the parameters in the model. In
particular, it is always possible to find a consistent model with any
desired occupation fraction in the range $0.1\leq f_{\rm occ}\leq
1.0$.  Nevertheless, in the models we can define three
different families of parameters based on their mass range $\Delta M
\equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ and occupation
fraction $f_{\rm occ}$. The dominant family is composed by models with
a narrow mass range $\Delta M<1.0$ dex, a low occupation fraction
$f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm  max}<
10^{12}$\hMsun.  The existence of this dominant family gives support
to the idea that the most massive dark matter halos at that epoch do
not host the brightest LAEs and that only a small fraction of star
forming galaxies can be actually detected as LAEs.

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations, markers of the galaxy formation process through
cosmic history \citep{ForeroRomero2012} and tracers of active star
formation  \citep{Guaita2013}

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter (DM), whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980}

Most models of galaxy formation find that the mass of the halo can be
used to predict properties of the galaxy such as its stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
mass.  For the reasons mentioned above finding the typical dark matter
halo mass hosting LAEs represents a significant step forward to
understand the nature of this population in the context of Lambda Cold
Dark Matter ($\Lambda$CDM) paradigm.

@@

Added to the uncertainties in the astrophysical processes describing
star formation in galactic populations, a highly debated step in this
approach is the calculation of the fraction of \ly photons
that escape the galaxy to the observer. Given the resonant nature of
the \lya line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM) \citep{Neufeld1991,ForeroRomero2011,
  Dijkstra2012,Laursen2013,Orsi2012}.

@@

Another approach uses the spatial information to infer the typical
mass of the hosting dark matter halo. This uses the fact that in cold
dark matter cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010}. In this
studies the observations were done on single fields of $\sim 1$
deg$^{2}$ and the conclusions derived on the halo host mass from
clustering information do not delve to deeply into the impact of
cosmic variance.

Recently \cite{Yamada2012} observed a wide area a bit more than ten
times the sky area covered so far, all of it with the same instrument
and the same criteria to reduce the observational data and produce LAE
catalogs. This large data-set allows a first observational study on the
expected cosmic variance effects in LAEs at $z=3.1$. This motivate us
to perform such a study from a theoretical persepective and study the
effects of the explicit inclusion of cosmica variance effects by using
a large N-body cosmological simulation. This
allows us to build mock fields resembling the observational geometry
and and directly expore the compatibility of a model with
observations.

In this \documentname, besides studying the impact of cosmic variance,
we implement a method to populate the DM halos with LAEs that bypasses
all the uncertainty involved in the estimation of star formation rates
and \ly escape fraction. Our model only defines whether a DM hosts a
detectable LAE or not. It does not predict a \lya luminosity. By using
this model we want to define the preferred mass range of DM halos
hosting these galaxies and the fraction of such halos that actually
can host a LAE.  Once the mock catalogs are constructed in this model
we proceed with a direct comparison against the statistics derived
from observations.

This \documentname is structured as follows. In the next section we present
the simulation and model used to produce the mock catalogs and the criteria
used to compare them against observations. In \SS \ref{sec:results} we
present the main results in terms of the parameters of our model. We
continue with a discussion of these results under the light of other
observational results for LAEs and abundance matching models for
galaxy formation before presenting our conclusions in \SS
\ref{sec:conclusions}.
